---
title: "SPRING '23 POA - Employment"
author: "Yinuo Zhou"
date: "7/27/2023"
output:
  pdf_document:
    toc: yes
    fig_width: 7
    fig_crop: no
  word_document:
    toc: yes
  html_document:
    toc: yes
    df_print: paged
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
source("/Users/zhouyinuo/Documents/communities-speak/repo/sp23/code/default_setup.R")
library(readr)
library(cli)
library(googledrive)
library(readxl)
library(tidyverse)
library(stringr)
library(haven)
library(rjson)
library(googlesheets4)
gs4_found <- gs4_find()
library(knitr)

target_census <- gs4_found %>% filter(str_detect(name, "target_demographics_sp23"))
wrangled <- read_csv("/Users/zhouyinuo/Documents/communities-speak/repo/sp23/data/output/wrangled.csv")
wrangled <- wrangled %>% mutate(gender = ifelse(gen == 1, "male", ifelse(gen == 2, "female", ifelse(gen == 3, "transmale", ifelse(gen == 4, "transfemale", ifelse(gen == 5, "queer", ifelse(gen == 5, "other", ifelse(gen == 6, "prefer not to say", NA))))))))

wrangled <- wrangled %>% mutate(hh_child = ifelse(hh_ch_0_17_bi == 0, "No children", ifelse(hh_ch_0_17_bi == 1, "Has children",NA)))

wrangled <- wrangled %>% mutate(hh_elderly = ifelse(hh_sn_65_bi == 0, "No senior citizens", ifelse(hh_sn_65_bi == 1, "Has senior citizens",NA)))

wrangled <- wrangled %>% mutate(inc_level = ifelse(inc_dist == 1, "HH income below $36500", ifelse(inc_dist == 2, "HH income between $36500 and $69500",ifelse(inc_dist == 3, "HH income above $69500",NA))))

wrangled <- wrangled %>% mutate(below_poverty = ifelse(below_poverty_bi == 0, "Above Poverty Level", ifelse(below_poverty_bi == 1, "Below Poverty Level",NA)))

wrangled <- wrangled %>% 
  mutate(below_median = ifelse(inc_be_med == 0, "Above Median Income", 
                         ifelse(inc_be_med == 1, "Below Median Income", NA)))
wrangled <- wrangled %>% 
  mutate(whole_pop = ifelse(resi_ny == 1, "New Yorker", NA))


race_target_census <- target_census %>% pull(id) %>%
  read_sheet(sheet = "Race")
race_weights <- race_target_census[!is.na(race_target_census$race_weight),
                                   c("label", "race_weight")] #%>%
 # mutate(across(race_weight, round, 5))

# only weighting black, white, hispanic, asian -- so chaning the others to 1
race_weights <- race_weights %>%
  rename(race_census = label) %>%
  filter(!grepl("indigenous", race_census)) %>%
  filter(!grepl("native", race_census)) %>%
  filter(!grepl("other", race_census)) %>%
  filter(!grepl("two", race_census))

# combining indigneous, native, other, and two into a single category "else"
wrangled <- wrangled %>%
  mutate(race_census = ifelse(grepl("indigenous", race_census), "else",
                              race_census)) %>%
  mutate(race_census = ifelse(grepl("native", race_census), "else",
                              race_census)) %>%
  mutate(race_census = ifelse(grepl("other", race_census), "else",
                              race_census)) %>%
  mutate(race_census = ifelse(grepl("two", race_census), "else",
                              race_census)) %>%
  mutate(race_census = ifelse(is.na(race_census), "else",
                              race_census))

if (!"else" %in% race_weights$race_census) {
  race_weights <- rbind(race_weights, data.frame(race_census = "else", race_weight = 1))
}



borough_target_census <- target_census %>% pull(id) %>%
  read_sheet(sheet = "Borough")
borough_weights <- borough_target_census[!is.na(borough_target_census$borough_weight),
                                   c("label", "borough_weight")] #%>%

borough_weights <- borough_weights %>%
  rename(borough = label)

wrangled <- wrangled %>%
  mutate(borough = ifelse(is.na(borough), "else",
                              borough))

if (!"else" %in% borough_weights$borough_census) {
  borough_weights <- rbind(borough_weights, data.frame(borough = "else", borough_weight = 1))
}



source("/Users/zhouyinuo/Documents/communities-speak/repo/sp23/reports/poa/make_tables.R")
source("/Users/zhouyinuo/Documents/communities-speak/repo/sp23/reports/poa/make_plots.R")
source("/Users/zhouyinuo/Documents/communities-speak/repo/sp23/reports/poa/calculate_weighted_percentages.R")
source("/Users/zhouyinuo/Documents/communities-speak/repo/sp23/reports/poa/calculate_weighted_nyc_race.R")
source("/Users/zhouyinuo/Documents/communities-speak/repo/sp23/reports/poa/calculate_weighted_nyc_boro.R")

```

# 1.1) People who changed employment status or employers in the last 6 months  [18&19]
```{r}


wrangled <- mutate(wrangled, emp_ch = as.numeric(emp_change))

#unweighted nyc as a whole 
df_na_rm <- filter(wrangled, !is.na(emp_ch))
(df_na_rm <- data.frame(count(df_na_rm,emp_ch)) %>% mutate(percent = round(100*n/sum(n), 3)))


#weighted nyc as a whole
weight_by_race<-weighted_percentages <- calculate_weighted_nyc(
  df = wrangled, 
  var_name = "emp_ch", 
  weights_df = race_weights, 
  weight_var = "race_weight", 
  label_true = "changed employment status or employers in the last 6 months", 
  label_false = "not changed employment status or employers in the last 6 months"
)

weight_by_race


weight_by_boro<- calculate_weighted_nyc_boro(
  df = wrangled, 
  var_name = "emp_ch", 
  weights_df = borough_weights, 
  weight_var = "borough_weight",
  label_true = "changed employment status or employers in the last 6 months", 
  label_false = "not changed employment status or employers in the last 6 months"
)
weight_by_boro

make_plots(wrangled, demographics, "emp_ch", 
                                title = "Proportion of People who changed employment status or employers in the last 6 months")

make_tables(wrangled, demographics, "emp_ch")

calculate_weighted_percentages(wrangled, demographics, "emp_ch", race_weights, "race_weight")


``` 


# 1.2) People who are currently unemployed and were employed 6 months ago [18 & 19]
```{r}
wrangled <- wrangled %>% mutate(emp_ch_unemp = ifelse(emp_before_hm != 1 & emp_before_st != 1 & emp_before_re != 1 & emp_ch == 1 & emp_after_un == 1, 1, 0))
#unweighted  nyc as a whole
df_na_rm <- filter(wrangled, !is.na(emp_ch_unemp))
(df_na_rm <- data.frame(count(df_na_rm,emp_ch_unemp)) %>% mutate(percent = round(100*n/sum(n), 3)))


#weighted nyc as a whole--race
weight_by_race<-weighted_percentages <- calculate_weighted_nyc(
  df = wrangled, 
  var_name = "emp_ch_unemp", 
  weights_df = race_weights, 
  weight_var = "race_weight", 
  label_true = "People who are currently unemployed and were employed 6 months ago", 
  label_false = "not the case"
)

weight_by_race

#weighted nyc as a whole--boro
weight_by_boro<- calculate_weighted_nyc_boro(
  df = wrangled, 
  var_name = "emp_ch_unemp", 
  weights_df = borough_weights, 
  weight_var = "borough_weight",
  label_true = "People who are currently unemployed and were employed 6 months ago", 
  label_false = "not the case"
)
weight_by_boro

make_plots(wrangled, demographics, "emp_ch_unemp", 
                                title = "Proportion of People who are currently unemployed and were employed 6 months ago")

make_tables(wrangled, demographics, "emp_ch_unemp")

calculate_weighted_percentages(wrangled, demographics, "emp_ch_unemp", race_weights, "race_weight")



```


# 1.3) Respondents whose income increased in the last 6 months were less likely to face job status changes in the last six months [17,18,19]
```{r}
#unweighted nyc as a whole- respondents whose income increased in the last 6 months
df_na_rm <- filter(wrangled, !is.na(inc_pos))
(df_na_rm <- data.frame(count(df_na_rm,inc_pos)) %>% mutate(percent = round(100*n/sum(n), 3))) %>% mutate(label = c("not the case ", "respondents whose income increased in the last 6 months"))


#weighted nyc as a whole--race-espondents whose income increased in the last 6 months
weight_by_race<-weighted_percentages <- calculate_weighted_nyc(
  df = wrangled, 
  var_name = "inc_pos", 
  weights_df = race_weights, 
  weight_var = "race_weight", 
  label_true = "respondents whose income increased in the last 6 months", 
  label_false = "not the case"
)

weight_by_race

#weighted nyc as a whole--boro-respondents whose income increased in the last 6 months
weight_by_boro<- calculate_weighted_nyc_boro(
  df = wrangled, 
  var_name = "inc_pos", 
  weights_df = borough_weights, 
  weight_var = "borough_weight",
  label_true = "respondents whose income increased in the last 6 months", 
  label_false = "not the case"
)
weight_by_boro


make_plots(wrangled, "inc_pos", "emp_ch", show = TRUE, title = "among people who got income increased, the percentage of them facing job status changes ")

make_plots(wrangled, "inc_neg", "emp_ch", show = TRUE, title = "among people who got income decreased, the percentage of them facing job status changes ")

make_plots(wrangled, "inc_same", "emp_ch", show = TRUE, title = "among people who got same income, the percentage of them facing job status changes ")

make_tables(wrangled, "inc_pos", "emp_ch")

make_tables(wrangled, "inc_neg", "emp_ch")
make_tables(wrangled, "inc_same", "emp_ch")

calculate_weighted_percentages(wrangled, "inc_pos", "emp_ch", race_weights, "race_weight")
calculate_weighted_percentages(wrangled, "inc_neg", "emp_ch", race_weights, "race_weight")
calculate_weighted_percentages(wrangled, "inc_same", "emp_ch", race_weights, "race_weight")
```



# 1.4) People with at least a Bachelor’s degree were more/less likely to face job status changes in the last six months [13,18,19]
```{r}
#unweighted nyc as a whole- proportion of respondents who have a Bachelor’s degree and above 

df_na_rm <- filter(wrangled, !is.na(sch_bach))
(df_na_rm <- data.frame(count(df_na_rm,sch_bach)) %>% mutate(percent = round(100*n/sum(n), 3)) %>% mutate(label = c("no bach", "bach")))

#weighted by race nyc as a whole--- proportion of respondents who have a Bachelor’s degree and above 
weight_by_race<-weighted_percentages <- calculate_weighted_nyc(
  df = wrangled, 
  var_name = "sch_bach", 
  weights_df = race_weights, 
  weight_var = "race_weight", 
  label_true = "proportion of respondents who have a Bachelor’s degree and above ", 
  label_false = "not the case"
)

weight_by_race

#weighted nyc as a whole--boro-proportion of respondents who have a Bachelor’s degree and above 
weight_by_boro<- calculate_weighted_nyc_boro(
  df = wrangled, 
  var_name = "sch_bach", 
  weights_df = borough_weights, 
  weight_var = "borough_weight",
  label_true = "proportion of respondents who have a Bachelor’s degree and above ", 
  label_false = "not the case"
)
weight_by_boro

make_plots(wrangled, "sch_bach", "emp_ch", show = TRUE, title = "the possibility of people with at least a Bachelor’s degree to face job status changes")
make_tables(wrangled, "sch_bach", "emp_ch")
calculate_weighted_percentages(wrangled, "inc_pos", "emp_ch", race_weights, "race_weight")
make_weighted_plots(wrangled, "inc_pos", "emp_ch", race_weights, "race_weight")

##unweighted nyc as a whole- People with at least a Bachelor’s degree were more/less likely to face job status changes in the last six months
wrangled <- wrangled %>% mutate(adverse = ifelse(inc_neg == 1 & emp_ch == 1, 1, 0))

make_plots(wrangled, "sch_bach", "adverse", show = TRUE, title = "the possibility of people with at least a Bachelor’s degree to face adverse job status changes")
make_tables(wrangled, "sch_bach", "adverse")
calculate_weighted_percentages(wrangled, "inc_pos", "adverse", race_weights, "race_weight")
make_weighted_plots(wrangled,"inc_pos", "adverse", race_weights, "race_weight")


```


# 1.5) People who are members of labor unions were less likely to face changes in job status [13,17,19]
```{r}
# unweighted nyc as a whole--- proportion of respondents who are members of labor unions 
wrangled <- wrangled %>% mutate(union_bi = ifelse(labor == 1, 1, 0))
df_na_rm <- filter(wrangled, !is.na(union_bi))
(df_na_rm <- data.frame(count(df_na_rm, union_bi)) %>% mutate(percent = round(100*n/sum(n), 3))) %>% mutate(label = c("non-labor union member", "labor union member"))

#weighted by race nyc as a whole--- proportion of respondents who are members of labor unions 
weight_by_race<-weighted_percentages <- calculate_weighted_nyc(
  df = wrangled, 
  var_name = "union_bi", 
  weights_df = race_weights, 
  weight_var = "race_weight", 
  label_true = " proportion of respondents who are members of labor unions ", 
  label_false = "not the case"
)

weight_by_race

#weighted nyc as a whole--boro- proportion of respondents who are members of labor unions
weight_by_boro<- calculate_weighted_nyc_boro(
  df = wrangled, 
  var_name = "union_bi", 
  weights_df = borough_weights, 
  weight_var = "borough_weight",
  label_true = " proportion of respondents who are members of labor unions", 
  label_false = "not the case"
)
weight_by_boro

make_plots(wrangled, "union_bi", "emp_ch", show = TRUE, title = "the possibility of people who are members of labor unions to face job status changes")
make_tables(wrangled, "union_bi", "emp_ch")

calculate_weighted_percentages(wrangled, "union_bi", "emp_ch", race_weights, "race_weight")
make_weighted_plots(wrangled,"union_bi", "emp_ch", race_weights, "race_weight")

```


#1.6)Persons whose income was at or above median income in 2022 were less likely to return to working in-person [16, 20]
```{r}
#unweighted data as a whole-respondents who are earning above median income
df_na_rm <- filter(wrangled, !is.na(inc_ab_med))
(df_na_rm <- data.frame(count(df_na_rm,inc_ab_med)) %>% mutate(percent = round(100*n/sum(n), 3))) %>% mutate(label = c("income at/below med", "income above med"))

#weighted by race nyc as a whole--- respondents who are earning above median income
weight_by_race<-weighted_percentages <- calculate_weighted_nyc(
  df = wrangled, 
  var_name = "inc_ab_med", 
  weights_df = race_weights, 
  weight_var = "race_weight", 
  label_true = " proportion of respondents who are earning above median income ", 
  label_false = "not the case"
)

weight_by_race

#weighted nyc as a whole--boro- proportion of respondents who are earning above median income
weight_by_boro<- calculate_weighted_nyc_boro(
  df = wrangled, 
  var_name = "inc_ab_med", 
  weights_df = borough_weights, 
  weight_var = "borough_weight",
  label_true = " proportion of respondents who are earning above median income", 
  label_false = "not the case"
)
weight_by_boro

#above
wrangled <- wrangled %>% mutate(hyb_rem = ifelse(wrk == 3 | wrk == 4, 1, 0))
make_tables(wrangled, "inc_ab_med","hyb_rem",  show = TRUE, title = "the possibility of people who earns above median income to beworking hybrid/remote")
make_plots(wrangled, "inc_ab_med","hyb_rem",  show = TRUE, title = "the possibility of people who earns above median income to beworking hybrid/remote")
calculate_weighted_percentages(wrangled, "inc_ab_med","hyb_rem", race_weights, "race_weight")
make_weighted_plots(wrangled,"inc_ab_med","hyb_rem", race_weights, "race_weight")

#below
make_plots(wrangled, "inc_be_med","hyb_rem",  show = TRUE, title = "the possibility of people who earns below median income to beworking hybrid/remote")

make_tables(wrangled, "inc_be_med","hyb_rem",  show = TRUE, title = "the possibility of people who earns below median income to beworking hybrid/remote")
calculate_weighted_percentages(wrangled, "inc_be_med","hyb_rem", race_weights, "race_weight")
make_weighted_plots(wrangled,"inc_be_med","hyb_rem", race_weights, "race_weight")

```


# 1.7) People who are unemployed and currently receiving unemployment benefits [19 & 22]
```{r}
wrangled <- wrangled %>% mutate(unem_benefits = ifelse(emp_after_un ==1 & unemp_ben==1, 1, 0))
#unweighted data-people who are unemployed and currently receiving unemployment benefits
df_na_rm <- filter(wrangled, !is.na(unem_benefits))
(df_na_rm <- data.frame(count(df_na_rm,unem_benefits)) %>% mutate(percent = round(100*n/sum(n), 3))) %>% mutate(label = "unemployed and receiving benefits")

#weighted by race nyc as a whole--- people who are unemployed and currently receiving unemployment benefits
weight_by_race<-weighted_percentages <- calculate_weighted_nyc(
  df = wrangled, 
  var_name = "unem_benefits", 
  weights_df = race_weights, 
  weight_var = "race_weight", 
  label_true = " proportion of people who are unemployed and currently receiving unemployment benefits ", 
  label_false = "not the case"
)

weight_by_race

#weighted nyc as a whole--boro- proportion of people who are unemployed and currently receiving unemployment benefits
weight_by_boro<- calculate_weighted_nyc_boro(
  df = wrangled, 
  var_name = "unem_benefits", 
  weights_df = borough_weights, 
  weight_var = "borough_weight",
  label_true = " people who are unemployed and currently receiving unemployment benefits", 
  label_false = "not the case"
)
weight_by_boro

make_tables(wrangled, demographics, "unem_benefits", show = TRUE, title = "people who are unemployed and currently receiving unemployment benefits")
make_plots(wrangled, demographics, "unem_benefits", show = TRUE, title = "unem_benefits")
calculate_weighted_percentages(wrangled, demographics, "unem_benefits", race_weights, "race_weight")
make_weighted_plots(wrangled, demographics, "unem_benefits", race_weights, "race_weight")

```


# 1.8)Reasons for employment change [21]

```{r}

emp_reas <- wrangled %>% select(starts_with("emp_reas_")) %>% colnames()
names(emp_reas) <- emp_reas

make_tables(wrangled, demographics,"emp_reas_laid")
make_plots(wrangled, demographics,"emp_reas_laid", title=  "laid off", show = TRUE)
calculate_weighted_percentages(wrangled, demographics, "emp_reas_laid", race_weights, "race_weight")
make_weighted_plots(wrangled, demographics, "emp_reas_laid", race_weights, "race_weight")

make_tables(wrangled, demographics,"emp_reas_fire")
make_plots(wrangled, demographics,"emp_reas_fire", title=  "fired", show = TRUE)
calculate_weighted_percentages(wrangled, demographics, "emp_reas_fire", race_weights, "race_weight")
make_weighted_plots(wrangled, demographics, "emp_reas_fire", race_weights, "race_weight")

make_tables(wrangled, demographics,"emp_reas_re")
make_plots(wrangled, demographics,"emp_reas_re", title=  "chose to retire", show = TRUE)
calculate_weighted_percentages(wrangled, demographics, "emp_reas_re", race_weights, "race_weight")
make_weighted_plots(wrangled, demographics, "emp_reas_re", race_weights, "race_weight")


make_tables(wrangled, demographics,"emp_reas_pay")
make_plots(wrangled, demographics,"emp_reas_pay", title=  "sought better pay", show = TRUE)
calculate_weighted_percentages(wrangled, demographics, "emp_reas_pay", race_weights, "race_weight")
make_weighted_plots(wrangled, demographics, "emp_reas_pay", race_weights, "race_weight")


make_tables(wrangled, demographics,"emp_reas_ben")
make_plots(wrangled, demographics,"emp_reas_ben", show = TRUE, title=  "sought better benefits")
calculate_weighted_percentages(wrangled, demographics, "emp_reas_ben", race_weights, "race_weight")
make_weighted_plots(wrangled, demographics, "emp_reas_ben", race_weights, "race_weight")


make_tables(wrangled, demographics,"emp_reas_covid")
make_plots(wrangled, demographics,"emp_reas_covid", show = TRUE, title=  "covid-19 concerns at my previous job")
calculate_weighted_percentages(wrangled, demographics, "emp_reas_covid", race_weights, "race_weight")
make_weighted_plots(wrangled, demographics, "emp_reas_covid", race_weights, "race_weight")



make_tables(wrangled,demographics, "emp_reas_new")
make_plots(wrangled,demographics, "emp_reas_new", show = TRUE, title=  "pivoted to a new industry/career")
calculate_weighted_percentages(wrangled, demographics, "emp_reas_new", race_weights, "race_weight")
make_weighted_plots(wrangled, demographics, "emp_reas_new", race_weights, "race_weight")


make_tables(wrangled,demographics, "emp_reas_train")
make_plots(wrangled,demographics, "emp_reas_train", show = TRUE, title=  "returned to school or job training")
calculate_weighted_percentages(wrangled, demographics, "emp_reas_train", race_weights, "race_weight")
make_weighted_plots(wrangled, demographics, "emp_reas_train", race_weights, "race_weight")


make_tables(wrangled,demographics, "emp_reas_other")
make_plots(wrangled,demographics, "emp_reas_other", show = TRUE, title=  "other")
calculate_weighted_percentages(wrangled, demographics, "emp_reas_other", race_weights, "race_weight")
make_weighted_plots(wrangled, demographics, "emp_reas_other", race_weights, "race_weight")


```


# 1.9) Reasons for difficulty in accessing unemployment benefits [23]
```{r}
unemp_diff <- wrangled %>% select(starts_with("unemp_diff_")) %>% colnames()
names(unemp_diff) <- unemp_diff

variables <- c("not_eng", "unemp_diff_lang", "unemp_diff_nav", "unemp_diff_app", 
               "unemp_diff_den", "unemp_diff_not_app", "unemp_diff_no_diff", 
               "unemp_diff_other", "unemp_diff_tech")

for (var in variables) {
    # 显示普通表格
    table_result <- make_tables(wrangled, demographics, var)
    print(table_result)
    
    # 显示普通条形图
    plot_result <- make_plots(wrangled, demographics, var, title = var, show = TRUE)
    print(plot_result)
    
    # 显示加权百分比表格
    weighted_percentages <- calculate_weighted_percentages(wrangled, demographics, var, race_weights, "race_weight")
    print(weighted_percentages)
    
    # 显示加权条形图
    weighted_plot <- make_weighted_plots(wrangled, demographics, var, race_weights, "race_weight")
    print(weighted_plot)
}



make_plots(wrangled, demographics, "not_eng", title=  "not_eng", show = TRUE)
make_plots(wrangled, demographics,"unemp_diff_lang", title=  "unemp_diff_lang", show = TRUE)
make_plots(wrangled, demographics,"unemp_diff_nav", title=  "unemp_diff_nav", show = TRUE)
make_plots(wrangled, demographics,"unemp_diff_app", title=  "unemp_diff_app", show = TRUE)
make_plots(wrangled, demographics,"unemp_diff_den", title=  "unemp_diff_den", show = TRUE)
make_plots(wrangled, demographics,"unemp_diff_not_app", show = TRUE, title=  "unemp_diff_not_app")
make_plots(wrangled, demographics,"unemp_diff_no_diff", show = TRUE, title=  "unemp_diff_none")
make_plots(wrangled,demographics, "unemp_diff_other", show = TRUE, title=  "unemp_diff_other")
make_plots(wrangled,demographics, "unemp_diff_tech", show = TRUE, title=  "unemp_diff_tech")

```





